Predicting computational reproducibility of data analysis pipelines \\in large population studies using collaborative filtering

### 2018-08-17 paper.tex
Minor fixes

\begin{abstract}
With the rise of the reproducibility crisis, evaluating the 
computational reproducibility of data analysis pipelines has become a 
critical issue. It is, however, a cumbersome 
process for analyses that involve data from large populations of 
subjects, due to their computational and storage requirements. 
We present a method to predict the computational 
reproducibility of data analysis pipelines in large population studies. 
We formulate the problem as a collaborative filtering process, with 
constraints on the construction of the training set. We propose 6 
different strategies to build the training set, which we evaluate on 2 
datasets, a synthetic one modeling a population with a growing number 
of subject types, and a real one obtained with neuroinformatics 
pipelines. Results show that one sampling method, ``Random File Numbers 
(Uniform)" is able to predict computational reproducibility with a good 
accuracy. We also analyze the relevance of including file and subject 
biases in the collaborative filtering model. We conclude that the 
proposed method is able to speed-up reproducibility 
evaluations substantially, with a reduced accuracy loss.
\end{abstract}

@@

Computational reproducibility, the ability to recompute data analyses 
over time and 
space~\cite{peng2011reproducible}, has become a critical component of 
scientific methodology as many researchers acknowledge the existence of 
a reproducibility crisis~\cite{baker2016there}. Infrastructural 
characteristics play an important role in making experiments 
reproducible, among other factors, such as data and code sharing. For 
instance, in 
neuroinformatics, our primary field of interest, several studies have 
shown the effect of the 
operating system on computational 
results~\cite{gronenschild2012effects, glatard2015reproducibility}.
However, conducting such reproducibility studies at scale is cumbersome 
due to the computational and storage requirements of data analysis 
pipelines.

Neuroinformatics pipelines are generally iterated on data coming from 10 to 
1,000 subjects, possibly with subtle input parameter variations to 
adjust specific data acquisition conditions. Subject data often capture 
anatomical and functional characteristics of their brain, for instance through
Magnetic Resonance Imaging (MRI) or Electroencephalography (EEG). The number of input files 
associated with a subject may vary, and these files may also be of 
different sizes. Typical data processing times range from 15 minutes to 
15 hours per subject, with input ranging from 100~MB to 15~GB per subject, and 
outputs ranging from 1~GB to 500~GB.

The reproducibility of a given pipeline may vary across subjects, for 
instance due to different pipeline branches being executed depending on 
the input data content. As an example, in 
the public database released by the Human Connectome 
Project~\cite{van2013wu}, subjects may have one or two anatomical 
images of each modality; when two images are present, they are
aligned together and averaged. Acquisition artifacts, for instance due to motion,
may also trigger corrections not otherwise required. Some 
branches of the pipeline may also be executed only for specific 
acquisition 
parameters. These remarks are 
consistent with recent findings that showed that variations in the 
results of a functional MRI analysis are dependent on the dataset being 
analyzed~\cite{bowring2018exploring}. To capture such variations, 
reproducibility evaluations need to be conducted on many subjects, 
which is unwieldy.

@@

Our goal is to predict the test set $\mathbb{T'}$ of missing values of 
$U$ from a training set $\mathbb{T}$ of known ones, where $\mathbb{T} 
\cap \mathbb{T'} = \emptyset$ and $\mathbb{T} \cup \mathbb{T'} = 
\{U_{i,j}\}$. In contrast with the traditional collaborative filtering 
problem, we have control over the construction of the training set, 
for instance we can choose to include only files producde by specific 
subjects, or the first files produced by the processing of every 
subject. Moreover, the construction of the training set is constrained by 
the 
order on the matrix rows, which is formalized as follows: 
\begin{equation}
\begin{array}{l}
\forall (i, j, k) \in \llbracket 1, N_f \rrbracket \times \llbracket 1, N_f \rrbracket \times \llbracket 1, N_s \rrbracket, \\
 \left( U_{i,k} \in \mathbb{T} \quad \mathrm{and} \quad U_{j,k} \in \mathbb{T'} \right) \Rightarrow i < j. \label{eq:time}
 \end{array}
\end{equation}

@@

\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.45\columnwidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_columns_04_training}
                  \caption{Complete Columns}
                  \label{fig:Columns-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\columnwidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_rows_04_training}
                  \caption{Complete Rows}
                  \label{fig:Rows-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\columnwidth}
                 \includegraphics[width=\columnwidth]{figures/5vs7_RS_04_training}
                \caption{Random Subjects (RS)}
                  \label{fig:Uniform-S-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\columnwidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_RFNU_04_training}
                  \caption{RFNU}
                  \label{fig:Diagonal-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\columnwidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_RFNT-L_04_training}
                  \caption{RFNTL}
                  \label{fig:triangular-L-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\columnwidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_RFNT-S_04_training}
                  \caption{RFNTS}
                  \label{fig:triangular-S-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{\columnwidth}
        \centering
        \includegraphics[width=0.45\columnwidth]{figures/5vs7_random-unreal_04_training}
                  \caption{Random Unreal}
                  \label{fig:Random-Unreal-Sample-Training-set}
        \end{subfigure}
        \caption{Training sets for different sampling methods; $\alpha=0.4$ (dark elements are in the training set).}
        \label{fig:sampling}
\end{figure}

\subsubsection{Random Subjects -- RS 
(Figure~\ref{fig:Uniform-S-Sample-Training-set})} This method builds 
the training set by selecting the files from random subjects until the 
training ratio is reached. The file selected in a subject is the file 
with the lowest index in this subject that has not been already 
selected in the training set. This respects the time constraints as files are 
sampled according to their modification time-stamps.

@@

          \textit{a} = (2\alpha - 1) N_{f} & \text{if $\alpha > 0.5$}
  \end{cases}
\]
 For $\alpha \leq 0.5$, we ensure that the average number of selected 
 files by subject is $\alpha N_f$, by sampling the number of 
 files in every subject from $U(0,N_f)$ with 
 probability $2\alpha$, and setting it to 0 otherwise. For $\alpha > 
 0.5$, this is ensured by the value of $a$, which leads the 
 average of U(a, b) to be $\alpha N_f$.

@@

As illustrated in Figures~\ref{fig:triangular-L-Sample-Training-set} 
and~\ref{fig:triangular-S-Sample-Training-set}, the motivation for the 
RFNTL method is to guarantee that, for large enough values of 
$\alpha$, all subjects will have at least a few files in the training 
set (no empty column), which is not the case for RFNU or RFNTS.
\begin{figure}
\centering
\includegraphics[width=0.5\columnwidth]{figures/triangular.pdf}
\caption{Triangular distribution $T(a, b, c)$}
\label{fig:triangular}
\end{figure}

\subsubsection{Random Unreal (Figure 
~\ref{fig:Random-Unreal-Sample-Training-set})} The training set is 
sampled in a random uniform way, regardless of the file generation 
times. This method does not respect the time constraints in 
Equation~\ref{eq:time}: it will be used as a baseline for comparison 
with other methods.

@@

We generated synthetic matrices as shown in 
Figure~\ref{fig:synthetic-data}. Each matrix has 100 files and 100 
subjects of different \emph{types}. Subjects of the same type behave 
identically and all types contain the same number of subjects $\pm$ 1. 
Such a decomposition by subject type correspond to the case where 
different subjects may have data of different nature, as is the case in the
dataset of the Human Connectome Project~\cite{van2013wu} where not all the 
subjects have the same amount of images.

 For subjects of a given type, the matrix consists of 
 $log(n)$  blocks, where $n$ is the number of types. Blocks are 
 defined with all the possible variation patterns: some types do not 
 vary at all, while other ones vary in every block. Such 
 patterns are meant to mimic the logic of data 
processing pipelines. Each block of files represents the files produced 
at a given stage of the pipeline, which may or may not contain 
reproducibility errors depending on the subject type.

@@

We collected data to evaluate the computational reproducibility of analysis
pipelines of the Human Connectome Project~\cite{glasser2013minimal}. We
processed a set S of 94 subjects randomly selected in the S500 HCP 
release\footnote{\url{https://db.humanconnectome.org}} in three 
execution conditions with different versions of the CentOS operating 
system (5.11, 6.8 and 7.2 -- referred as C5, C6 and C7), using the 
PreFreesurfer and Freesurfer pipelines described 
in~\cite{glasser2013minimal} and available on 
GitHub\footnote{\url{https://github.com/Washington-University/Pipelines/releases/tag/v3.19.0}}. 
For each pipeline, we identified the set F of files produced for all
subjects in all conditions. For each condition pair and each pipeline, 
we computed a binary reproducibility matrix U of size $|F|\times|S|$, 
where $U_{i,j}$ is true if and only if file $i$ of subject $j$ was different in 
each condition. Rows of $U$ were ordered by ascending file modification 
time in a random subject in S.

@@

\section{Results}
\label{sec:results}
We conducted two experiments for each reproducibility matrix, to evaluate the performance of our 
predictions using (1) ALS without biases, and (2) ALS with subject and file biases. We 
compare the performance of our sampling methods to (1) a dummy 
classifier that always predicts the value in the majority class and (2) 
the Random Unreal method, used as the baseline sampling technique. All 
reported values are averages over 5 repetitions. Due to sampling 
issues, it is possible that the actual training ratio obtained with 
some of the sampling methods does not 
match the target one. We checked that the difference between the target 
and real training ratios was lower than 0.01. We used Spark's ALS model 
as available in package \texttt{pyspark.ml}, with 50 factors, a 
regularization parameter of 0.01, a maximum of 5 iterations and non 
negative constraints set to true. The code and data used to obtain the results are available through GitHub at
\url{https://github.com/big-data-lab-team/paper-reproducibility-collaborative-filtering}.
\todo{Soudabeh: push your code to reprotools. Tristan: release and check.}
\todo{publish code on zenodo}

@@

\subsubsection{ALS without Bias}
 Figure~\ref{fig:results-synthetic} shows accuracy results for ALS 
 without bias. By construction, the accuracy of the dummy classifier is 
 close to 0.5 for all subject types. Random Unreal performs well 
 for all subject types, which confirms that ALS is working correctly. 
 All the other methods 
perform better than the dummy classifier, and their accuracy decreases 
as the number of subject types increases.
However, only 3 methods provide accuracy values above 0.85 
for all subject types: Random Subjects (RS), RFNU and RFNTL. 
Surprisingly, RFNTS does not perform well for more than 2 subject 
types, even for $\alpha=0.9$.

@@

\subsection{Effect of the maximum number of iterations}
Figure~\ref{fig:iterations} illustrates the effect of the number of 
iterations used in the ALS optimization, for the 
RFNU method and the synthetic dataset with 8 types. 
From a training ratio of 0.7, the number of iterations has a moderate 
effect on the accuracy. We used 5 iterations in our experiments, 
which slightly degrades the accuracy compared to 15 or 20.

@@

Figure~\ref{fig:overlap} compares the prediction errors made by RFNU, 
RFNTL and RFNTS. The training set is represented
in black (negatives) and white (positives), and the test set is in green 
(true positives), yellow (false negatives), gray (true negatives) and 
red (false positives). This representation provides insights regarding 
where, and perhaps why, prediction errors 
occur. At this training ratio, RFNU (Figure~\ref{fig:overlap-rfnu}) 
uniformly samples the number of files per subject between 0.8$N_f$ and 
$N_f$, which enables the training on the last 
files of the pipeline, while maintaining a low number of columns with 
a low training ratio. On the contrary, RFNTL 
(Figure~\ref{fig:overlap-rfntl}) samples the number of files per 
subject from 0.85$N_f$, but the probability to have a complete column 
is 0 (the one complete column in Figure~\ref{fig:overlap-rfntl} is the 
one included to prevent cold start issues), which leads to a ``stripe" 
of prediction errors at the bottom of the matrix. RFNTS 
(Figure~\ref{fig:overlap-rfnts}) doesn't have this issue, as it 
includes many complete columns in the training set. However, it comes at 
the cost of several columns with a number of files lower than 
60\%: in such columns, the prediction is often entirely wrong, which explains
the reduced accuracy compared to RFNU.

Figure~\ref{fig:error-locality} shows the results obtained by RFNU on 
the Freesurfer dataset for a training ratio of 0.6. The prediction 
error is localized (1) at the bottom of the matrix, which corresponds 
to the end of the execution, and (2) in the regions where file 
reproducibility varies across subjects, i.e., lines are not entirely 
black or entirely white.

@@

From a practical standpoint, our study shows that reproducibility 
evaluations of the PreFreesurfer and Freesurfer pipelines can be 
conducted using only 50\% of the files produced, with an accuracy above 
95\%. Potentially, this could reduce the computing time and storage 
required for such studies by a factor of 2. However, such studies could 
not be conducted by processing only half of the subjects entirely, 
which would correspond to the Complete Columns sampling method. 
Instead, the processing of all the subjects should be initiated, and it 
should be terminated in a random uniform way, assuming that files are 
uniformly produced throughout the execution. On more complex matrices, 
such as the synthetic dataset studied here, the training ratio required 
to get a 95\% accuracy increases to 0.85.

% What are the limitations
Our study could be extended to real-valued reproducibility 
matrices instead of just binary ones. It is indeed common for file 
differences to be quantified using specific similarity measures or 
distances, such as Levenshtein distance between strings, or the sum of 
squared distances between voxels of an image. Our sampling methods 
could be directly applied to real-valued matrices, and we expect our 
conclusions regarding the best-performing sampling method (RFNU) and 
the inclusion of biases in the model to remain valid.

>>>>

\begin{abstract}
Evaluating the 
computational reproducibility of data analysis pipelines has become a 
critical issue. It is, however, a cumbersome 
process for analyses that involve data from large populations of 
subjects, due to their computational and storage requirements. 
We present a method to predict the computational 
reproducibility of data analysis pipelines in large population studies. 
We formulate the problem as a collaborative filtering process, with 
constraints on the construction of the training set. We propose 6 
different strategies to build the training set, which we evaluate on 2 
datasets, a synthetic one modeling a population with a growing number 
of subject types, and a real one obtained with neuroinformatics 
pipelines. Results show that one sampling method, ``Random File Numbers 
(Uniform)" is able to predict computational reproducibility with a good 
accuracy. We also analyze the relevance of including file and subject 
biases in the collaborative filtering model. We conclude that the 
proposed method is able to speed-up reproducibility 
evaluations substantially, with a reduced accuracy loss.
\end{abstract}

@@

Computational reproducibility, the ability to recompute analyses 
over time and 
space~\cite{peng2011reproducible}, has become a critical component of 
scientific methodology as many researchers acknowledge the existence of 
a reproducibility crisis~\cite{baker2016there}. Among other factors, infrastructural 
characteristics play an important role in making experiments 
reproducible. For 
instance, in 
neuroinformatics, our primary field of interest, studies have 
shown the effect of the 
operating system on computational 
results~\cite{gronenschild2012effects, glatard2015reproducibility}.
However, conducting such reproducibility studies at scale is cumbersome 
due to the computational and storage requirements of analysis 
pipelines.

Neuroinformatics pipelines are generally iterated on data coming from 10 to 
1,000 subjects, possibly with subtle input parameter variations to 
adjust specific data acquisition conditions. Subject data often capture 
anatomical and functional characteristics of their brain, for instance through
Magnetic Resonance Imaging (MRI) or Electroencephalography (EEG). The number of input files 
associated with a subject may vary, and these files may also be of 
different sizes. Typical processing times range from 15 minutes to 
15 hours per subject, with input ranging from 100~MB to 15~GB per subject, and 
outputs ranging from 1~GB to 500~GB.

The reproducibility of a given pipeline may vary across subjects, for 
instance due to different pipeline branches being executed depending on 
the input data content. As an example, in 
the public database released by the Human Connectome 
Project~\cite{van2013wu}, subjects may have one or two anatomical 
images of each modality; when two images are present, they are
aligned together and averaged. Acquisition artifacts, for instance due to motion,
may also trigger corrections not otherwise required, and some 
branches of the pipeline may be executed only for specific 
acquisition 
parameters. These remarks are 
consistent with recent findings showing that variations in the 
results of a functional MRI analysis are dependent on the dataset being 
analyzed~\cite{bowring2018exploring}. To capture such variations, 
reproducibility evaluations need to be conducted on many subjects, 
which is unwieldy.

@@

Our goal is to predict the test set $\mathbb{T'}$ of missing values of 
$U$ from a training set $\mathbb{T}$ of known ones, where $\mathbb{T} 
\cap \mathbb{T'} = \emptyset$ and $\mathbb{T} \cup \mathbb{T'} = 
\{U_{i,j}\}$. In contrast with the traditional collaborative filtering 
problem, we have control over the construction of the training set. 
For instance, we can choose to include only files produced by specific 
subjects, or the first files produced by the processing of every 
subject. Moreover, the construction of the training set is constrained by 
the 
order of the matrix rows, which is formalized as follows: 
\begin{equation}
\begin{array}{l}
\forall (i, j, k) \in \llbracket 1, N_f \rrbracket \times \llbracket 1, N_f \rrbracket \times \llbracket 1, N_s \rrbracket, \\
 \left( U_{i,k} \in \mathbb{T} \quad \mathrm{and} \quad U_{j,k} \in \mathbb{T'} \right) \Rightarrow i < j. \label{eq:time}
 \end{array}
\end{equation}

@@

\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.45\columnwidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_columns_04_training}
                  \caption{Complete Columns}
                  \label{fig:Columns-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\columnwidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_rows_04_training}
                  \caption{Complete Rows}
                  \label{fig:Rows-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\columnwidth}
                 \includegraphics[width=\columnwidth]{figures/5vs7_RS_04_training}
                \caption{Random Subjects (RS)}
                  \label{fig:Uniform-S-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\columnwidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_RFNU_04_training}
                  \caption{RFNU}
                  \label{fig:Diagonal-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\columnwidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_RFNT-L_04_training}
                  \caption{RFNTL}
                  \label{fig:triangular-L-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\columnwidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_RFNT-S_04_training}
                  \caption{RFNTS}
                  \label{fig:triangular-S-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{\columnwidth}
        \centering
        \includegraphics[width=0.45\columnwidth]{figures/5vs7_random-unreal_04_training}
                  \caption{Random Unreal}
                  \label{fig:Random-Unreal-Sample-Training-set}
        \end{subfigure}
        \caption{Training sets for different sampling methods, $\alpha=0.4$ (dark elements are in the training set).}
        \label{fig:sampling}
\end{figure}

\subsubsection{Random Subjects -- RS 
(Figure~\ref{fig:Uniform-S-Sample-Training-set})} This method builds 
the training set by selecting the files from random subjects until the 
training ratio is reached. The file selected in a subject is the file 
with the lowest index in this subject that has not been already 
selected in the training set, which respects the time constraints.

@@

          \textit{a} = (2\alpha - 1) N_{f} & \text{if $\alpha > 0.5$}
  \end{cases}
\]
 For $\alpha \leq 0.5$, we ensure that the average number of selected 
 files by subject is $\alpha N_f$ by sampling the number of 
 files in every subject from $U(0,N_f)$ with 
 probability $2\alpha$, and setting it to 0 otherwise. For $\alpha > 
 0.5$, this is ensured by the value of $a$, which leads the 
 average of U(a, b) to be $\alpha N_f$.

@@

As illustrated in Figures~\ref{fig:triangular-L-Sample-Training-set} 
and~\ref{fig:triangular-S-Sample-Training-set}, the point of the 
RFNTL method is to guarantee that, for large enough values of 
$\alpha$, all subjects will have at least a few files in the training 
set (no empty column), which is not the case for RFNU or RFNTS.
\begin{figure}
\centering
\includegraphics[width=0.5\columnwidth]{figures/triangular.pdf}
\caption{Triangular distribution $T(a, b, c)$}
\label{fig:triangular}
\end{figure}

\subsubsection{Random Unreal (Figure 
~\ref{fig:Random-Unreal-Sample-Training-set})} The training set is 
sampled in a random uniform way, regardless of the file creation 
times. This method does not respect the time constraints in 
Equation~\ref{eq:time}: it will be used as a baseline for comparison.

@@

We generated synthetic matrices as shown in 
Figure~\ref{fig:synthetic-data}. Each matrix has 100 files and 100 
subjects of different \emph{types}. Subjects of the same type behave 
identically and all types contain the same number of subjects $\pm$ 1. 
Such a decomposition by subject type corresponds to a situation where 
different subjects may have data of different nature, as is the case in the
Human Connectome Project data~\cite{van2013wu} where not all the 
subjects have the same amount of images.

 For subjects of a given type, the matrix consists of 
 $log(n)$  blocks, where $n$ is the number of types. Blocks are 
 defined with all the possible variation patterns: some types do not 
 vary at all, while other ones vary between every block. Such 
 patterns are meant to mimic the logic of data 
processing pipelines: each block of files represents the files produced 
at a given stage of the pipeline, which may or may not contain 
reproducibility errors depending on the subject type.

@@

We
processed a set S of 94 subjects randomly selected in the S500 HCP 
release\footnote{\url{https://db.humanconnectome.org}} of the Human 
Connectome Project~\cite{glasser2013minimal}, in three execution 
conditions with different versions of the CentOS operating system 
(5.11, 6.8 and 7.2 -- referred as C5, C6 and C7), using the 
PreFreesurfer and Freesurfer pipelines described 
in~\cite{glasser2013minimal} and available on 
GitHub\footnote{\url{https://github.com/Washington-University/Pipelines/releases/tag/v3.19.0}}. 
For each pipeline, we identified the set F of files produced for all 
subjects in all conditions. For each condition pair and each pipeline, 
we computed a binary reproducibility matrix U of size $|F|\times|S|$, 
where $U_{i,j}$ is true if and only if file $i$ of subject $j$ was 
different in each condition. Rows of $U$ were ordered by ascending file 
modification time in a random subject in S.

@@

\section{Results}
\label{sec:results}
We conducted two experiments for each reproducibility matrix, to evaluate the performance of our 
predictions using (1) ALS without biases, and (2) ALS with subject and file biases. We 
compare the performance of our sampling methods to (1) a dummy 
classifier that always predicts the value in the majority class and (2) 
the Random Unreal method, used as the baseline sampling technique. All 
reported values are averages over 5 repetitions. Due to sampling 
issues, it is possible that the actual training ratio obtained with 
some of the sampling methods does not exactly
match the target one. We checked that the difference between the target 
and real training ratios was lower than 0.01. We used Spark's ALS model 
as available in \texttt{pyspark.ml}, with 50 factors, a 
regularization parameter of 0.01, a maximum of 5 iterations and non 
negative constraints set to true. The code and data used to obtain the results are available through GitHub at
\url{https://github.com/big-data-lab-team/paper-reproducibility-collaborative-filtering}.
\todo{Soudabeh: push your code to reprotools. Tristan: release and check.}
\todo{publish code on zenodo}

@@

\subsubsection{ALS without Bias}
 Figure~\ref{fig:results-synthetic} shows accuracy results for ALS 
 without bias. By construction, the accuracy of the dummy classifier is 
 close to 0.5 for all subject types. Random Unreal performs well 
 for all subject types, which confirms that ALS is working correctly. 
 All the other methods 
perform better than the dummy classifier, and their accuracy decreases 
as the number of subject types increases.
However, only 3 methods can provide accuracy values above 0.85 
for all subject types: Random Subjects (RS), RFNU and RFNTL. 
Surprisingly, RFNTS does not perform well for more than 2 subject 
types, even for $\alpha=0.9$.

@@

\subsection{Effect of the maximum number of iterations}
Figure~\ref{fig:iterations} illustrates the effect of the number of 
iterations used in the ALS optimization, for the 
RFNU method and the synthetic dataset with 8 types. 
From a training ratio of 0.7, the number of iterations has a moderate 
effect on the accuracy. We used 5 iterations in our experiments, 
which only slightly degrades the accuracy compared to 15 or 20.

@@

Figure~\ref{fig:overlap} compares the prediction errors made by RFNU, 
RFNTL and RFNTS. The training set is represented
in black (negatives) and white (positives), and the test set is in green 
(true positives), yellow (false negatives), gray (true negatives) and 
red (false positives). This representation provides insights regarding 
where, and perhaps why, prediction errors 
occur. At this training ratio, RFNU (Figure~\ref{fig:overlap-rfnu}) 
uniformly samples the number of files per subject between 0.8$N_f$ and 
$N_f$, which enables the training on the last 
files of the pipeline, while maintaining a low number of columns with 
a low training ratio. On the contrary, RFNTL 
(Figure~\ref{fig:overlap-rfntl}) samples the number of files per 
subject from 0.85$N_f$, but the probability to have a complete column 
is 0 (the one complete column in Figure~\ref{fig:overlap-rfntl} is the 
one included to prevent cold start issues), which leads to a ``stripe" 
of prediction errors at the bottom of the matrix. RFNTS 
(Figure~\ref{fig:overlap-rfnts}) does not have this issue, as it 
includes many complete columns in the training set. However, it comes at 
the cost of several columns with a number of files lower than 
60\%: in such columns, the prediction is often entirely wrong, which explains
the reduced accuracy compared to RFNU.

Figure~\ref{fig:error-locality} shows the RFNU results on 
the Freesurfer dataset for a training ratio of 0.6. The prediction 
error is localized (1) at the bottom of the matrix, which corresponds 
to the end of the execution, and (2) in the regions where file 
reproducibility varies across subjects, i.e., lines are not entirely 
black or entirely white. This is consistent with our expectations and 
confirms the validity of our results.

@@

From a practical standpoint, our study shows that reproducibility 
evaluations of the PreFreesurfer and Freesurfer pipelines can be 
conducted using only 50\% of the files produced, with an accuracy above 
95\%. Potentially, this could reduce the computing time and storage 
required for such studies by a factor of 2. However, such studies could 
not be conducted by processing only half of the subjects entirely, 
which would correspond to the complete columns sampling method. 
Instead, the processing of all the subjects should be initiated and 
terminated in a random uniform way, assuming that files are uniformly 
produced throughout the execution. On more complex matrices, such as 
the synthetic dataset studied here, the training ratio required 
to get a 95\% accuracy increases to 0.85.

% What are the limitations
Our study could be extended to real-valued reproducibility 
matrices instead of just binary ones. It is indeed common for file 
differences to be quantified using specific similarity measures or 
distances, such as the Levenshtein distance between strings, or the sum of 
squared distances among voxels of an image. Our sampling methods 
could be directly applied to real-valued matrices, and we expect our 
conclusions on the best-performing sampling method (RFNU) and 
the inclusion of biases in the model to remain valid.
